### 2022-05-27 main/chapter_2ljets.tex
flesh out btagging

\begin{figure}[tp]
\centering
\includegraphics[width=\textwidth]{figures/2ljets_summary_log.pdf}
\caption{%
Data of the $\twoljets$ electroweak analysis with \emph{post-fit}
backgrounds~\cite{atlas2022searches}.
The lower panel shows $S_\mathrm{ATLAS}$ from
Equation~\ref{eqn:significance_atlas}.
Control, validation and signal regions are shown from left to right, with the
regions within each category ordered approximately by their typical $\met$.
Likelihoods from validation regions are not included in the fit.
The `Top' category contains $t\bar t$ and $tW$ processes, and
`Other' contains fake/non-prompt lepton, higgs, triboson, $t\bar tZ$, and other
rare top processes.%
}
\label{fig:2ljets_summary}
\end{figure}

@@

Tagging adds labels to objects.
Unlike how electrons and muons are very cleanly separated however, jet tagging
is a more noisy affair.

For jets in $\twoljets$, $b$-tagging is used to identify jets which are
more likely to have originated from $b$ quarks;
these tags serve to improve our separation of $t\bar t$ backgrounds
(since their weak decays almost always produce $b$ quarks), and
of signals producing $Z/h \rightarrow b\bar b$ decays in the GMSB model.

Signal $b$-tagged jets require $\pt > 20\,\eV[G]$, $|\eta| < 2.5$ and a tag from
the `MV2c10' algorithm at its $77\%$ efficiency working point.
This tagging algorithm uses
\cite{atlas_jet_mv2c10_2017, atlas_jet_tagging_2019}
\TODO{cite, summarize}
%
boosted decision tree~\cite{friedman2002stochastic}

encode information derived from data within the jet cluster.
In particular, $b$-tagging attempts to identify jets originating from $b$
quarks, which have distinct features due to the long lifetimes and decay
channels of $B$ hadrons.

\TODO{context: charm quarks, large radius (jumbo) jets top higgs}

@@

In the context of $\twoljets$ selections, $\metsig$ suppresses $t\bar t$
contributions more than $\met$ alone, plausibly since $t\bar t$ events are
accompanied with more hadronic activity than other sources of $\ptmiss$.
What remains of the standard model in the high-$\metsig$ tail for $\twoljets$
selections is almost pure diboson -- $ZZ$ with a hard $Z\rightarrow \nu\nu$,
and $W^\pm Z$ where the charged lepton in $W^\pm\rightarrow\ell^\pm\nu$
is not reconstructed.
\TODO{plot of diboson decomposition here?}
These effects are demonstrated from simulations in
Figure~\ref{fig:2ljets_presel_met}.

\begin{figure}[tp]
\centering
\includegraphics[width=0.49\textwidth]{figures/2ljets_presel_met_logy.png}
\hfill
\includegraphics[width=0.49\textwidth]{figures/2ljets_presel_met_sig_logy.png}
\caption{
Illustration of how $\metsig$ beats $\met$ alone in sensitivity to
example signal models.
At large $\metsig$, $t\bar t$ and other backgrounds are suppressed, leaving
quite pure diboson backgrounds and greater significance measures.
Signal models with smaller mass splittings tend to have the bulk of their
events at smaller $\met$.
\\[0.5em]
This selection requires two same-flavor opposite-sign signal leptons with
$\pt^{\ell_2} > 25\,\eV[G]$,
$\mll \in (71, 111)\,\eV[G]$,
$\mjj \in (60, 110)\,\eV[G]$,
$\nbtag \leq 1$,
$\met > 100\,\eV[G]$, and
$\metsig > 6$.
\\[0.5em]
``Significance'' displayed in the lower subplot uses the Binomial significance
$Z_{Bi}$ from~\cite{cousins2008evaluation}, with a $30\%$ background
uncertainty and yields taken to the right of a given value.
Large values indicate expected sensitivity to the signal model.
}
\label{fig:2ljets_presel_met}
\end{figure}

@@

Lacking direct observation of sleptons, $\mttwo$ remains useful in separating
standard model processes; dileptonic $t\bar t$ and $W^\pm W^\mp$ decays both
satisfy $\mttwo(\ell^+, \ell^-, 0) \leq m(W)$ (since $0 \approx m(\nu)$).
The $\twoljets$ electroweak search depends on this effect;
most selections require $\mttwo(\ell^+, \ell^-, 0) > 80\,\eV[G]$, or similar,
to reduce $t\bar t$, $W^\pm W^\mp$ and $\tau^\pm\tau^\mp$ backgrounds.

>>>>

\begin{figure}[tp]
\centering
\includegraphics[width=\textwidth]{figures/2ljets_summary_log.pdf}
\caption{%
Data of the $\twoljets$ electroweak analysis with \emph{post-fit}
backgrounds~\cite{atlas2022searches}.
The lower panel shows $S_\mathrm{ATLAS}$ from
Equation~\ref{eqn:significance_atlas}.
Control, validation and signal regions are shown from left to right, with the
regions within each category ordered approximately by their typical $\met$.
Likelihoods from validation regions are not included in the fit.
The `Top' category contains $\ttbar$ and $tW$ processes, and
`Other' contains fake/non-prompt lepton, higgs, triboson, $\ttbar Z$, and other
rare top processes.%
}
\label{fig:2ljets_summary}
\end{figure}

@@

Tagging adds labels to objects.
Unlike how electrons and muons are very cleanly separated however, jet tagging
is a more noisy affair.
Taggers are available for various jet properties, such charm flavor ($c$-jet),
bottom flavor ($b$-jet).
Of these, $\twoljets$ uses only $b$-tagging.

Larger radius jets can also use taggers for heavy decays, such as those from
$W$, Higgs or top resonance, but large radius were considered too much
additional work to considered in $\twoljets$.

Jets originating from $b$ quarks are distinct due to the long lifetimes and
long decay trees of $B$-hadrons, both of which leave signals in \atlas'
inner trackers.

For jets in $\twoljets$, $b$-tagging is useful to separate events due to
$\ttbar$ backgrounds (since their weak decays almost always produce $b$
quarks), and to select for signals producing $Z/h \rightarrow b\bar b$ decays
in the GMSB model.

Signal $b$-jets in $\twoljets$ require $\pt > 20\,\eV[G]$, $|\eta| < 2.5$ and a
tag from the `MV2c10' boosted decision tree
algorithm~\cite{
friedman2002stochastic,
atlas_jet_mv2c10_2017,
atlas_jet_tagging_2019
} at its $77\%$ efficiency working point.
This efficiency means that a true $b$-jet sampled from its testing set has a
$77\%$ probability of being tagged.
It is trained on a mixture of $\ttbar$ and $\zjets$ simulated data,
and the `c10' suffix means that those data include a $7\%$ fraction of
$c$-jets~\cite{atlas_twiki_mv2}.

This tagging algorithm combines various lower-level features, including the
basic kinematics $\pt$ and $|\eta|$, and more intricate features of
observable secondary vertices including reconstructed hadronic decay
trees~\cite{atlas_jet_mv2c10_2017}.

Alternative working points of $70\%$ and $85\%$ efficiency are used in similar
\atlas\ searches.
These were trialled and found to make negligible differences to $\twoljets$.

An alternative $b$-tagging algorithm named `DL1' is now recommended by \atlas,
but its performance is ``similar'' to MV2, and the two tag
``highly correlated'' samples~\cite{atlas_jet_tagging_2019}, so we did not
attempt to switch.

@@

In the context of $\twoljets$ selections, $\metsig$ suppresses $\ttbar$
contributions more than $\met$ alone, plausibly since $\ttbar$ events are
accompanied with more hadronic activity than other sources of $\ptmiss$.
What remains of the standard model in the high-$\metsig$ tail for $\twoljets$
selections is almost pure diboson -- $ZZ$ with a hard $Z\rightarrow \nu\nu$,
and $W^\pm Z$ where the charged lepton in $W^\pm\rightarrow\ell^\pm\nu$
is not reconstructed.
\TODO{plot of diboson decomposition here?}
These effects are demonstrated from simulations in
Figure~\ref{fig:2ljets_presel_met}.

\begin{figure}[tp]
\centering
\includegraphics[width=0.49\textwidth]{figures/2ljets_presel_met_logy.png}
\hfill
\includegraphics[width=0.49\textwidth]{figures/2ljets_presel_met_sig_logy.png}
\caption{
Illustration of how $\metsig$ beats $\met$ alone in sensitivity to
example signal models.
At large $\metsig$, $\ttbar$ and other backgrounds are suppressed, leaving
quite pure diboson backgrounds and greater significance measures.
Signal models with smaller mass splittings tend to have the bulk of their
events at smaller $\met$.
\\[0.5em]
This selection requires two same-flavor opposite-sign signal leptons with
$\pt^{\ell_2} > 25\,\eV[G]$,
$\mll \in (71, 111)\,\eV[G]$,
$\mjj \in (60, 110)\,\eV[G]$,
$\nbtag \leq 1$,
$\met > 100\,\eV[G]$, and
$\metsig > 6$.
\\[0.5em]
``Significance'' displayed in the lower subplot uses the Binomial significance
$Z_{Bi}$ from~\cite{cousins2008evaluation}, with a $30\%$ background
uncertainty and yields taken to the right of a given value.
Large values indicate expected sensitivity to the signal model.
}
\label{fig:2ljets_presel_met}
\end{figure}

@@

Lacking direct observation of sleptons, $\mttwo$ remains useful in separating
standard model processes; dileptonic $\ttbar$ and $W^\pm W^\mp$ decays both
satisfy $\mttwo(\ell^+, \ell^-, 0) \leq m(W)$ (since $0 \approx m(\nu)$).
The $\twoljets$ electroweak search depends on this effect;
most selections require $\mttwo(\ell^+, \ell^-, 0) > 80\,\eV[G]$, or similar,
to reduce $\ttbar$, $W^\pm W^\mp$ and $\tau^\pm\tau^\mp$ backgrounds.
